---
title: "Homework 1"
format: html
editor: visual
author: "Megan Brown"
---

# Problem 1

## Problem 1(a)

```{r}

# column names from abalone.names
columns <- c("sex", "length", "diameter", "height", "whole_weight",
             "shucked_weight", "viscera_weight", "shell_weight", "rings")

# data from abalone.data
df <- read.csv('datasets/abalone/abalone.data', header=FALSE)

# set new column names
names(df) <- columns
```

## Problem 1(b)

```{r}

# count the number of records by sex 
table(df$sex)
```

There are 1,307 females, 1,342 infants, and 1,528 males.

## Problem 1(c)

### Problem 1(c)(1)

I found the R function for correlation from [this blog post](https://statsandr.com/blog/correlation-coefficient-and-correlation-test-in-r/).

```{r}

# correlation between different weights and rings
print(paste0("Whole Weight: ", cor(df$whole_weight, df$rings)))

print(paste0("Shucked Weight: ", cor(df$shucked_weight, df$rings)))

print(paste0("Viscera Weight: ", cor(df$viscera_weight, df$rings)))

print(paste0("Shell Weight: ", cor(df$shell_weight, df$rings)))

```

Shell weight is most correlated with rings.

### Problem 1(c)(2)

```{r}
# split dataframes by sex
males <- df[df$sex == "M",]
females <- df[df$sex == "F",]
infants <- df[df$sex == "I",]

# output correlations
print(paste0("Males: ", cor(males$shell_weight, males$rings)))
print(paste0("Females: ", cor(females$shell_weight, females$rings)))
print(paste0("Infants: ", cor(infants$shell_weight, infants$rings)))

```

For infants, shell weight was most correlated with rings.

### Problem 1(c)(3)

```{r}
# get the record where the number of rings is the highest number of rings
df[df$rings == max(df$rings),]
```

For the abalone with the most rings, the whole weight was 1.8075g, the shucked weight was 0.7055g, the viscera weight was 0.3215g, and the shell weight was 0.475g.

### Problem 1(c)(4)

```{r}
numerator <- nrow(df[df$viscera_weight > df$shell_weight,]) 
denominator <- nrow(df)

print(100*numerator/denominator)
```

The percentage of abalones for which the viscera weight was greater than the shell weight was 6.51%.
